Railway: Adaptive Disk Storage for Interaction Graphs

### 2014-09-11 docs/edbt2015/text/introduction.tex
minor progress over intro

>>>>

We are living in an ever more connected world, where the data generated by
people, software systems, and the physical world is more accessible than
before and is much larger in volume, variety, and velocity. In many
application domains, such as telecommunications and social media, live data
recording the interactions between people, systems, and the environment is
available for analysis. This data often takes the form of a temporally
evolving graph, where entities are the vertices and the interactions between
them are the edges. We call such graphs \emph{interaction graphs}.

Data analytics performed on interaction graphs can bring new business insights
and improve decision making. For instance, the graph structure may represent
the interactions in a social network, where finding communities in the graph
can facilitate targeted advertising. In the telecommunications (telco) domain,
call details records (CDRs) can be used to capture the call interactions
between people, and locating closely connected groups of people can be used
for generating promotions.

Interaction graphs are temporal in nature, and more importantly, they are
append-only. This is in contrast to relationship graphs, which are updated via
insertion and deletion operations. An example of a relationship graph is a
social network capturing the follower-followee relationship among users.
Examples of interactions graphs include CDR graphs capturing calls between
telco customers or mention graphs capturing interactions between users of a
micro-blogging service. The append-only nature of the interaction graphs make
storing them on disk a necessity. Furthermore, the analysis of this historical
interaction data form an important aspect of the analytical landscape.

The ability to efficiently support historical analysis over interaction graphs
require effective solutions for the problem of \emph{data layout} on the disk.

% Importance of the data layout on disk and locality 
% Earlier work on blocks formed of temporal neighborlists

% Attributes associated with interactions
% I/O inefficiency due to attributes that are not accessed
% Similar to the problem of row vs column stores

% Challenges: finding the right partitioning
% what if the access pattern changes over time
% query access patterns are different over different time ranges

% Solution: railway layout 
% adaptive layout optimization
% algorithms for optimal partitioning
% heuristics & evaluation
